---
title: "1.4 Portfolio 1 Jake Arlow"
output: html_notebook
---



```{r}
library("BIEN")
library(ape) #Package for working with phylogenies in R
library(maps) #Useful for making quick maps of occurrences
library(sp) # A package for spatial data
library(dplyr)
library(lattice)
```

Alright here's the occurence and range map for American Chestnut: 

```{r}
chestnut_range <- BIEN_ranges_load_species("Castanea dentata")
chestnut_occurrence <- BIEN_occurrence_species(species = "Castanea dentata")
map('world', fill=T, col="grey", bg="light blue", xlim=c(-180,-20), ylim=c(-60, 80))
plot(chestnut_range, col="green", add=T)
points(cbind(chestnut_occurrence$longitude, chestnut_occurrence$latitude), col="blue", pch=20, cex=1)
```
There's chestnuts in Hawaii? Wild.

Next: trait data 

```{r}
Castanea_dentata <- BIEN_trait_traitbyspecies(trait = "whole plant height", species = "Castanea dentata")
str(Castanea_dentata)
Castanea_dentata
```

Mean whole plant height for american chestnut:

```{r}
chestnut = Castanea_dentata[c(3)] 
chestnut = as.data.frame(chestnut)
chestnut_height = as.numeric(chestnut$trait_value)
mean(chestnut_height, data=chestnut)
```

alright now i'm gonna compare it with Chinese chestnut:

```{r}
Castanea_mollissima <- BIEN_trait_traitbyspecies(trait = "whole plant height", species = "Castanea mollissima")
str(Castanea_crenata)
Castanea_mollissima
```

Mean whole plant height for Chinese chestnut:

```{r}
C_chestnut = Castanea_mollissima[c(3)] 
C_chestnut = as.data.frame(C_chestnut)
C_chestnut_height = as.numeric(C_chestnut$trait_value)
mean(C_chestnut_height, data=C_chestnut)
```

making a box plot for American Chestnut:

```{r}
boxplot(chestnut_height)
```

Box plot for Chinese Chestnut:

```{r}
boxplot(C_chestnut_height)
```

I'm not sure if this really worked... (the Chinese chestnut box plot)

T-test for the two chestnut species:

```{r}
t.test(chestnut_height, C_chestnut_height)
```

Now I'm gonna do this again but for breadfruit (trying it again bc it didn't work for me when we did it in lab):


Range and occurence map for Artocarpus altilis:

```{r}
breadfruit_range <- BIEN_ranges_load_species("Artocarpus altilis")
breadfruit_occurrence <- BIEN_occurrence_species(species = "Artocarpus altilis")
map('world', fill=T, col="grey", bg="light blue", xlim=c(-180,-20), ylim=c(-60, 80))
plot(breadfruit_range, col="green", add=T)
points(cbind(breadfruit_occurrence$longitude, breadfruit_occurrence$latitude), col="blue", pch=20, cex=1)
```


Trait data for Artocarpus altilis:

```{r}
breadfruit1 <- BIEN_trait_traitbyspecies(trait = "whole plant height", species = "Artocarpus altilis")
str(breadfruit1)
breadfruit1
```


Trait data for other species of breadfruit:

```{r}
breadfruit2 <- BIEN_trait_traitbyspecies(trait = "whole plant height", species = "Artocarpus hirsutus")
str(breadfruit2)
breadfruit2
```


mean whole plant height Artocarpus altilis:

```{r}
breadfruit1 = breadfruit1[c(3)] 
breadfruit1 = as.data.frame(breadfruit1)
breadfruit1_height = as.numeric(breadfruit1$trait_value)
mean(breadfruit1_height, data=breadfruit1)
```


mean whole plant height Artocarpus hirsutus:

```{r}
breadfruit2 = breadfruit2[c(3)] 
breadfruit2 = as.data.frame(breadfruit2)
breadfruit2_height = as.numeric(breadfruit2$trait_value)
mean(breadfruit2_height, data=breadfruit2)
```


making a box plot for Artocarpus altilis:

```{r}
boxplot(breadfruit1_height)
```

making a box plot for Artocarpus hirsutus:

```{r}
boxplot(breadfruit2_height)
```


T-test for the two breadfruit species:

```{r}
t.test(breadfruit1_height, breadfruit2_height)
```

# Awesome! You seem to have R figured out pretty well. The only thing that didn't work was line 49 `str(Castanea_crenata)`, but I think it's because you just forgot to rename "crenata" to "mollissima". I hope you had fun with the assignment, or at least found it useful!